A Multi-telescope Campaign on \frb: Implications for the FRB Population

### 2017-05-22 repeater-vladata.tex
shriharsh comments in

\begin{abstract}
We present results of the coordinated observing campaign that made the first subarcsecond localization of a Fast Radio Burst, \frb. During this campaign, we made the first simultaneous detection of an FRB burst by multiple telescopes: the VLA at 3~GHz and the Arecibo Observatory at 1.4~GHz. Of the nine bursts detected by the Very Large Array at 3~GHz, four had simultaneous observing coverage at other observatories. We use multi-observatory constraints and modeling of bursts seen only at 3~GHz to confirm earlier results showing that burst spectra are not well modeled by a power law. We find that burst spectra are characterized by a $\sim500$~MHz envelope and apparent radio energy as high as $10^{40}$\ erg. We measure significant changes in the apparent dispersion between bursts that can be attributed to frequency-dependent profiles or some other intrinsic burst structure that adds a systematic error to the estimate of DM by up to 1\%. We use \frb\ as a prototype of the FRB class to estimate a volumetric birth rate of FRB sources $R_{\rm{FRB}} \approx 5\times10^{-5}/N_r$\ Mpc$^{-3}$\ yr$^{-1}$, where $N_r$\ is the number of bursts per source over its lifetime. This rate is broadly consistent with models of FRBs from young pulsars or magnetars born in superluminous supernovae or long gamma-ray bursts, if the typical FRB repeats on the order of $10^3$\ times during its lifetime.
\end{abstract}

@@

This paper is part of a series based on the first localization of an FRB and its unambiguous association to an extragalactic host \citep{LOC, OPT, EVN}. \frb, also known as the ``repeating FRB'', was discovered \citep{2014ApJ...790..101S} in data acquired in 2012 as part of the PALFA survey of the Arecibo Observatory \citep{2006ApJ...637..446C, lazarus}. In mid 2015, new Arecibo observations revealed a series of bursts at a similar DM and sky position that rules out cataclysmic models for this source \citep{2016Natur.531..202S}. The typical DM measured by early Arecibo observations was 559 pc cm$^{-3}$ \citep{2016arXiv160308880S}, although somewhat higher values have been seen in more recent observations (560.5 pc cm$^{-3}$; Hessels et al., in prep).

@@

Observations in August and September were searched by a prototype version of \rf\footnote{See \url{http://realfast.io} and \citep{2017AAS...22933002L}.}. \rf\ is a real-time, fast imaging transient search system. The current prototype runs on CPU-based hardware that is normally dedicated to the VLA correlator; for this experiment, it runs the transient search pipeline software called \emph{rtpipe}\footnote{See \url{https://github.com/caseyjlaw/rtpipe} and \citet{2015ApJ...807...16L}.}. Images were formed for each integration with DMs of 0, 546, 556.9, 560, and 565 pc cm$^{-3}$ and a time resampling grid of 5, 10, 20, 40, and 80 ms. This DM grid was chosen to maintain 90\% sensitivity to the nominal DM range of 540--570 pc cm$^{-3}$. Gain calibration was made from observations of J0555+3948 by an automated system (telcal), which uses phase-only calibration. A flux scale was calculated for each spectral window from an observation of 3C48 and applied to all burst spectra and has an accuracy of about 10\%.

@@

Table \ref{tab:spec} summarizes the spectrotemporal properties of all nine VLA bursts. Parameters such as integrated flux density, S/N, and Stokes V were measured from the burst properties integrated in frequency. Stokes V values were estimated from the dual-circular feeds as $(RR-LL)/(RR+LL)\approx3\%$, which is comparable to the systematic error expected from beam squint at this location in the primary beam \footnote{See EVLA memo \#195 at \url{https://library.nrao.edu/public/memos/evla/EVLAM_195.pdf}.}. Given that systematic effects dominate the apparent circular polarization, this observation limits the fractional circular polarization to less than 3\%.

@@

 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.5\columnwidth]{spec_57649.pdf}
  \includegraphics[width=0.5\columnwidth]{specAO_57648.pdf}
 \end{minipage}
\caption{Ten panels show spectra for nine VLA bursts and one Arecibo burst (bottom right). The VLA spectra are drawn from flux-calibrated, dedispersed 5~ms integrations. The Arecibo spectrum is drawn from a 2.7~ms window with 3.125~MHz channels and the flux scale is esimated using the radiometer equation. The solid line shown with VLA spectra is a best-fit Gaussian model found through modeling. The typical VLA flux density error per channel is 70~mJy, which is shown at the top right of each panel next to the MJD label.
\label{fig:spec}}
\end{center}
\end{figure*}

@@

Assuming that the burst detection probability follows a Poisson distribution, the mean VLA, 3~GHz burst rate is $R=0.16\pm0.05\ \rm{hr}^{-1}$ above a fluence of 0.2 Jy ms. The nondetection in the first half of 3~GHz observing limits the FRB rate to $R<0.1$\ hour$^{-1}$\ (95\% confidence limit). The 3~GHz burst rate was much higher during the late-2016 campaign, $R=0.3\pm0.1$\ hour$^{-1}$.

The integrated burst flux density is calculated by integrating the burst spectral model (Table \ref{tab:spec}) in frequency and time. Some VLA burst spectra seem to be contained by the 2.5 to 3.5~GHz band and most of them seem to have Gaussian envelopes that are well modeled by the emission within that band. Assuming that the Gaussian shape defines the full emission window, the integrated 3~GHz flux density can be converted to a total isotropic energy with no further assumptions about the burst spectral properties as:
\begin{equation}
E_{\rm{int}} = (\rm{A}\ 5\ \rm{ms}) ( \sigma_{ch}\ 2.355 4\ \rm{MHz}) 10^{-23} (4\pi L_d^2)\ \rm{erg}
\end{equation}
\noindent where the first term represents the burst fluence, the second term represents the integral of the burst spectrum, 2.355 scales $\sigma$\ to FWHM, and $L_d$\ is the luminosity distance to the source.

Figure \ref{fig:ed} shows the \frb\ burst energy cumulative distribution as seen by the VLA and calculated from prior observations by Arecibo and the Green Bank Telescope \citep{2016Natur.531..202S,2016arXiv160308880S}. The latter two energy distributions are scaled from the fluence by assuming that all burst energy is included by the observation. This likely underestimates the burst energy for some bursts, but should not affect the slope of the distribution. The VLA distribution represents a relatively long campaign, so it is sensitive to lower event rates, but is less sensitive than the single-dish campaigns shown (minimum $E_{\rm{int}}\approx3\times10^{38}$\ erg). We also show the rate upper limit (95\% confidence) from the early-2016 VLA campaign to demonstrate that even identical observing campaigns have different detection rates.

@@

We present the first simultaneous detection of an FRB with multiple telescopes and over frequencies from 1.2 to 3.5~GHz. The flux density of the Arecibo burst detected on MJD 57648 is an order of magnitude less than that seen by the VLA. At the same time, three other bursts from \frb\ had similar observing coverage but were not detected simultaneously. This is consistent with the spectral structure observed within the 3~GHz VLA band, which is typically limited to a Gaussian envelope with width of roughly 500~MHz. This confirms earlier results at 1.4~GHz \citep{2016Natur.531..202S,2016arXiv160308880S} with a wider bandwidth and extends the presence of this phenomenon to 3~GHz.

@@

After correcting for barycentric and dispersion delays, the VLA+AO burst spectrum has a residual temporal drift. This drift can be interpreted as an excess DM, but contemporaneous observations at 1.4~GHz measured a significantly lower DM (560.5 pc cm$^{-3}$; Hessels et al, in prep). This excess DM is comparable to the burst-to-burst variation in DM within the whole sample of VLA 3~GHz bursts. This variation had been noted in earlier 1.4~GHz Arecibo observations of \frb\ \citep{2016Natur.531..202S
, 2016arXiv160308880S}, but the VLA bursts demonstrate that this burst-specific DM is seen simultaneously from 1.4 to 3~GHz.

@@

The clustering of burst times for the whole FRB population has been modeled as a ``red spectrum'' \citep{2016MNRAS.458L..89C}, while recent work has modeled \frb\ burst times with a modified Poisson distribution \citep{oppermann}. Both temporal models imply that existing observational constraints on repetition are weaker than expected from Poissonian statistics \citep{2015MNRAS.454..457P, 2015ApJ...807...16L}. If the clustered bursts are typical of the FRB population, then wide, shallow surveys are the preferred strategy for blind detection of FRBs. Also, since ``bursts predict bursts'', many short observations are more likely to detect an FRB than a single long observation of the same total length and all new FRB detections should be immediately and intensively followed up.

@@

Despite its many assumptions, this volumetric birth rate is more appropriate for comparison to rates used for other classes of transient. The LGRB and SLSN-I rates are $10^{-7}$\ and $10^{-8}$\ Mpc$^{-3}$\ yr$^{-1}$, respectively \citep{2007ApJ...657L..73G,2012Sci...337..927G}. Comparing the FRB rate to these rates implies that FRB-emitting sources can be generated in LGRBs and SLSN-I if FRBs repeat $5\times10^2$\ and $5\times10^3$\ times\footnote{\citet{2017arXiv170400022N} provide a similar estimate of FRB birth rate and reach a similar conclusion.}, respectively. This assumes isotropic burst energies of $\sim10^{40}$~erg or intrinsic energies of $\sim10^{39} (\Omega_b/0.1)$~erg. \citet{2017arXiv170102370M} use the young magnetar model for FRBs to calculate the maximum number of bursts that can be powered by its magnetic field as **$\approx3000 (0.1/\Omega_b)$. This rate comparison suggests that the SLSN-I rate is most consistent with the FRB rate, although there is significant room for adjustment in the models and rate estimate.

@@

% observatories
The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc..
Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration.
% systems
This research made use of Astropy, a community-developed core Python package for Astronomy (Astropy Collaboration, 2013).
%individuals
CJL is supported by the University of California Office of the President under Lab Fees Research Program Award 237863 and NSF award 1611606. MAM is supported by NSF award 1458952. 
JWTH, CGB, and DM acknowledge support from the European Research Council under the European Union's Seventh Framework Programme (FP/2007-2013) / ERC Grant Agreement nr.\ 337062 (DRAGNET; PI Hessels). JWTH also acknowledges funding from an NWO Vidi fellowship.
KPM's research is supported by the Oxford Centre for Astrophysical Surveys which is funded through the Hintze Family Charitable Foundation. AS gratefully acknowledges support from the European Research Council under grant ERC-2012- StG-307215 LODESTONE. The AMI-LA telescope gratefully acknowledges support from the European Research Council under grant ERC-2012- StG-307215 LODESTONE, the UK Science and Technology Facilities Council (STFC) and the University of Cambridge. 
PS holds a Covington Fellowship at DRAO. 
The LWA1 station is supported by the National Science Foundation under grant 1139974 of the University Radio Observatory program.
VMK acknowledges support from NSERC, CIFAR, the Canada Research Chair Program and from the Lorne Trottier Chair in Astrophysics \& Cosmology.

>>>>

\begin{abstract}
We present results of the coordinated observing campaign that made the first subarcsecond localization of a Fast Radio Burst, \frb. During this campaign, we made the first simultaneous detection of an FRB burst by multiple telescopes: the VLA at 3~GHz and the Arecibo Observatory at 1.4~GHz. Of the nine bursts detected by the Very Large Array at 3~GHz, four had simultaneous observing coverage at other observatories. We use multi-observatory constraints and modeling of bursts seen only at 3~GHz to confirm earlier results showing that burst spectra are not well modeled by a power law. We find that burst spectra are characterized by a $\sim500$~MHz envelope and apparent radio energy as high as $10^{40}$\ erg. We measure significant changes in the apparent dispersion between bursts that can be attributed to frequency-dependent profiles or some other intrinsic burst structure that adds a systematic error to the estimate of DM by up to 1\%. We use \frb\ as a prototype of the FRB class to estimate a volumetric birth rate of FRB sources $R_{\rm{FRB}} \approx 5\times10^{-5}/N_r$\ Mpc$^{-3}$\ yr$^{-1}$, where $N_r$\ is the number of bursts per source over its lifetime. This rate is broadly consistent with models of FRBs from young pulsars or magnetars born in superluminous supernovae or long gamma-ray bursts, if the typical FRB repeats on the order of thousands of times during its lifetime.
\end{abstract}

@@

This paper is part of a series based on the first localization of an FRB and its unambiguous association to an extragalactic host \citep{LOC, OPT, EVN}. \frb, also known as the ``repeating FRB'', was discovered \citep{2014ApJ...790..101S} in data acquired in 2012 as part of the PALFA survey of the Arecibo Observatory \citep{2006ApJ...637..446C, 2015ApJ...812...81L}. In mid 2015, new Arecibo observations revealed a series of bursts at a similar DM and sky position that rules out cataclysmic models for this source \citep{2016Natur.531..202S}. The typical DM measured by early Arecibo observations was 559 pc cm$^{-3}$ \citep{2016arXiv160308880S}, although somewhat higher values have been seen in more recent observations (560.5 pc cm$^{-3}$; Hessels et al., in prep).

@@

Observations in August and September were searched by a prototype version of \rf\footnote{See \url{http://realfast.io} \citep{2017AAS...22933002L}.}. \rf\ is a real-time, fast imaging transient search system. The current prototype runs on CPU-based hardware that is normally dedicated to the VLA correlator; for this experiment, it runs the transient search pipeline software called \emph{rtpipe}\footnote{See \url{https://github.com/caseyjlaw/rtpipe} \citep{2015ApJ...807...16L}.}. Images were formed for each integration with DMs of 0, 546, 556.9, 560, and 565 pc cm$^{-3}$ and a time resampling grid of 5, 10, 20, 40, and 80 ms. This DM grid was chosen to maintain 90\% sensitivity to the nominal DM range of 540--570 pc cm$^{-3}$. Gain calibration was made from observations of J0555+3948 by an automated system (telcal), which uses phase-only calibration. A flux scale was calculated for each spectral window from an observation of 3C48 and applied to all burst spectra and has an accuracy of about 10\%.

@@

Table \ref{tab:spec} summarizes the spectrotemporal properties of all nine VLA bursts. Parameters such as integrated flux density, S/N, and Stokes V were measured from the burst properties integrated in frequency. The circular polarization fraction (Stokes V/I) was estimated as $(RR-LL)/(RR+LL)\approx3\%$, which is comparable to the systematic error expected from beam squint at this location in the primary beam \footnote{See EVLA memo \#195 at \url{https://library.nrao.edu/public/memos/evla/EVLAM_195.pdf}.}. Given that systematic effects dominate the apparent circular polarization, this observation limits the fractional circular polarization to less than 3\%.

@@

 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.5\columnwidth]{spec_57649.pdf}
  \includegraphics[width=0.5\columnwidth]{specAO_57648.pdf}
 \end{minipage}
\caption{Ten panels show spectra for nine VLA bursts and one Arecibo burst (bottom right). The VLA spectra are drawn from flux-calibrated, dedispersed 5~ms integrations. The Arecibo spectrum is drawn from a 2.7~ms window with 3.125~MHz channels and the flux scale is estimated using the radiometer equation. The solid line shown with VLA spectra is a best-fit Gaussian model found through modeling; no comparable modeling is done for the Arecibo spectrum. The typical VLA flux density error per channel is 70~mJy, which is shown at the top right of each panel next to the MJD label.
\label{fig:spec}}
\end{center}
\end{figure*}

@@

Assuming that the burst detection probability follows a Poisson distribution, the mean VLA, 3~GHz burst rate is $R=0.16\pm0.05\ \rm{hr}^{-1}$ above a fluence of 0.2 Jy ms. The nondetection in the first half of 3~GHz observing limits the FRB rate to $R<0.1$\ hour$^{-1}$\ (95\% confidence limit). The 3~GHz burst rate was much higher during the late-2016 campaign, $R=0.3\pm0.1$\ hour$^{-1}$. This confirms recent analysis of published bursts from \frb\ \citep{2017arXiv170504881O}.

The integrated burst flux density is calculated by integrating the burst spectral model (Table \ref{tab:spec}) in frequency and time. Some VLA burst spectra seem to be contained by the 2.5 to 3.5~GHz band and most of them seem to have Gaussian envelopes that are well modeled by the emission within that band. Assuming that the Gaussian shape defines the full emission window, the integrated 3~GHz flux density can be converted to a total isotropic energy with no further assumptions about the burst spectral properties as:
\begin{equation}
E_{\rm{int}} = (\rm{A} \times 5\ \rm{ms}) ( 2.355 \sigma_{ch}\ \times 4\ \rm{MHz}) 10^{-23} (4\pi L_d^2)\ \rm{erg}
\end{equation}
\noindent where the first term represents the burst fluence, the second term represents the integral of the burst spectrum, 2.355 scales $\sigma$\ to FWHM, and $L_d$\ is the luminosity distance to the source.

Figure \ref{fig:ed} shows the \frb\ burst energy cumulative distribution as seen by the VLA and calculated from prior observations by Arecibo and the Green Bank Telescope \citep{2016Natur.531..202S, 2016arXiv160308880S}. The latter two energy distributions are scaled from the fluence by assuming that all burst energy is included by the observation. This likely underestimates the burst energy for some bursts, but should not affect the slope of the distribution. The VLA distribution represents a relatively long campaign, so it is sensitive to lower event rates, but is less sensitive than the single-dish campaigns shown (minimum $E_{\rm{int}}\approx3\times10^{38}$\ erg). We also show the rate upper limit (95\% confidence) from the early-2016 VLA campaign to demonstrate that even identical observing campaigns have different detection rates.

@@

We present the first simultaneous detection of an FRB with multiple telescopes and over frequencies from 1.2 to 3.5~GHz. The flux density of the Arecibo burst detected on MJD 57648 is an order of magnitude less than that seen by the VLA. At the same time, three other bursts from \frb\ had similar observing coverage but were not detected simultaneously. This is consistent with the spectral structure observed within the 3~GHz VLA band, which is typically limited to a Gaussian envelope with width of roughly 500~MHz. This confirms earlier results at 1.4~GHz \citep{2016Natur.531..202S, 2016arXiv160308880S} with a wider bandwidth and extends the presence of this phenomenon to 3~GHz.

@@

After correcting for barycentric and dispersion delays, the VLA+AO burst spectrum has a residual temporal drift. This drift can be interpreted as an excess DM, but contemporaneous observations at 1.4~GHz measured a significantly lower DM (560.5 pc cm$^{-3}$; Hessels et al, in prep). This excess DM is comparable to the burst-to-burst variation in DM within the whole sample of VLA 3~GHz bursts. This variation had been noted in earlier 1.4~GHz Arecibo observations of \frb\ \citep{2016Natur.531..202S,2016arXiv160308880S}, but the VLA bursts demonstrate that this burst-specific DM is seen simultaneously from 1.4 to 3~GHz.

@@

The clustering of burst times for the whole FRB population has been modeled as a ``red spectrum'' \citep{2016MNRAS.458L..89C}, while recent work has modeled \frb\ burst times with a modified Poisson distribution \citep{2017arXiv170504881O}. Both temporal models imply that existing observational constraints on repetition are weaker than expected from Poissonian statistics \citep{2015MNRAS.454..457P, 2015ApJ...807...16L}. If the clustered bursts are typical of the FRB population, then wide, shallow surveys are the preferred strategy for blind detection of FRBs. Also, since ``bursts predict bursts'', many short observations are more likely to detect an FRB than a single long observation of the same total length and all new FRB detections should be immediately and intensively followed up.

@@

Despite its many assumptions, this volumetric birth rate is more appropriate for comparison to rates used for other classes of transient. The LGRB and SLSN-I rates are $10^{-7}$\ and $10^{-8}$\ Mpc$^{-3}$\ yr$^{-1}$, respectively \citep{2007ApJ...657L..73G,2012Sci...337..927G}. Comparing the FRB rate to these rates implies that FRB-emitting sources can be generated in LGRBs and SLSN-I if FRBs repeat $5\times10^2$\ and $5\times10^3$\ times\footnote{\citet{2017arXiv170400022N} provide a similar estimate of FRB birth rate and reach a similar conclusion.}, respectively. This assumes isotropic burst energies of $\sim10^{40}$~erg or intrinsic energies of $\sim10^{39} (\Omega_b/0.1)$~erg. \citet{2017arXiv170102370M} use the young magnetar model for FRBs to calculate the maximum number of bursts that can be powered by its magnetic field as $\approx3000 (0.1/\Omega_b)$. This rate comparison suggests that the SLSN-I rate is most consistent with the FRB rate, although there is significant room for adjustment in the models and rate estimate.

@@

% observatories
The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc..
Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration.
% systems
This research made use of Astropy, a community-developed core Python package for Astronomy (Astropy Collaboration, 2013).
%individuals
CJL is supported by the University of California Office of the President under Lab Fees Research Program Award 237863 and NSF award 1611606. MAM is supported by NSF award 1458952. 
JWTH, CGB, and DM acknowledge support from the European Research Council under the European Union's Seventh Framework Programme (FP/2007-2013) / ERC Grant Agreement nr.\ 337062 (DRAGNET; PI Hessels). JWTH also acknowledges funding from an NWO Vidi fellowship.
SPT acknowledges support from the McGill Astrophysics Fellowship.
KPM's research is supported by the Oxford Centre for Astrophysical Surveys which is funded through the Hintze Family Charitable Foundation. AS gratefully acknowledges support from the European Research Council under grant ERC-2012- StG-307215 LODESTONE. The AMI-LA telescope gratefully acknowledges support from the European Research Council under grant ERC-2012- StG-307215 LODESTONE, the UK Science and Technology Facilities Council (STFC) and the University of Cambridge. 
PS holds a Covington Fellowship at DRAO. 
The LWA1 station is supported by the National Science Foundation under grant 1139974 of the University Radio Observatory program.
VMK acknowledges support from NSERC, CIFAR, the Canada Research Chair Program and from the Lorne Trottier Chair in Astrophysics \& Cosmology.
